---
title: "Case Study: Kernel Density Hotspots"
author: "Michael Branion-Calles"
date: "March 23, 2020"
output: 
  html_document:
    code_folding: hide 
    fig_width: 7
    fig_height: 7
    toc: true
    toc_float:
      theme: 
      collapsed: false
      smooth_scroll: false
editor_options: 
  chunk_output_type: console
---

# Background

Hotspot in traffic safety context is a location where there are more unsafe incidents occurring than is normal. 

- The road system can be conceptualised as consisting of three components: the environment, the road users and their vehicles (Haddon, 1980; Sayed et al., 1995).

- Different factors within these components will interact in space and time to contribute causally to a given crash that occurs in the system. Spatial patterns of crashes that have occurred within a broader region and time period are therefore realizations of underlying spatial processes, in which crash risk factors related to road environment, road user and vehicles are spatially heterogeneous.

- The areas in which there are frequently occurring crashes represent areas of high crash burden, whereas risky areas are place where there are a higher probability of a crash occurring, relative to the number of opportunities (i.e. road users interacting in that space) for a crash to occur (Fuller and Morency, 2013). Measurements of the number of opportunities for a crash to occur is often referred to as “exposure” and are a fundamental data requirement in analysis of risk. Without exposure data, spatial analysis of where crashes occur will identify areas with a higher probability of a crash, but necessarily a higher risk, since it will not control for the amount of traffic.

- Here we define hotspots as a location where the probability or risk of a crash is much higher than the rest of the study area. 

- The detection of spatial hotspots is both the first step in understanding the spatial processes that contribute towards risky or burdensome locations in a jurisdiction, and a simple method for defining problem areas with minimal data requirements.

- In road safety, there is a distinction between crash burden and crash risk. 
  - The first aim of this paper is to review commonly used methods for detecting spatial hotspots in road safety, including kernel density estimators and local measures of spatial association. 
  - The second aim of this paper is to apply these techniques to a case study of bicycling safety data and compare the results. 


# Case Study Data

- IBIMS region as study area 
- Strava Metro Counts as Exposure
- BikeMaps near misses and collisions as incidents
  - ICBC left out due to snapping of incidents to nearest intersection or midblock, gives a potentially very misleading view of the spatial distribution of crashes, which will become more problematic with network kernel density techniques

```{r, warning=FALSE,message=FALSE,results=FALSE}
library(tidyr)
library(raster)
library(spatstat)
library(maptools)
library(sf)
library(dplyr)
library(ggplot2)
library(mapview)

#load functions
file.sources = list.files(path = paste0(getwd(),"/R/functions/"),pattern="*.R$",full.names = TRUE)
sapply(file.sources,source,.GlobalEnv)

study_area <- read_sf(paste0(getwd(),"/input/processed/study_area.gpkg"))

network <- read_sf(paste0(getwd(),"/input/processed/edge_ec_201601_201709_total.gpkg"))

incidents <- read_sf(paste0(getwd(),"/input/processed/inc_bm_201601_201709_snp_30m.gpkg")) %>% 
  filter(st_intersects(x = ., y = st_as_sfc(st_bbox(study_area)), sparse = FALSE))


ggplot() +
  geom_sf(data = network,color="lightgrey",alpha=0.05)+
  geom_sf(data = incidents,aes(color=p_type),show.legend = "point")+
  scale_color_brewer(name = "Incident Type",
                     type = "qual")+
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.minor = element_blank() ,
        panel.grid.major = element_blank() ,
        axis.title = element_blank(),
        legend.position = "left") + ggtitle("Bicycling Incidents in the Victoria Region: \nJanuary 2016 - September 2017")

mapview(network,z="CAADB") + mapview(incidents, z= "p_type")

```

# Planar Kernel Density Surfaces

Kernel Density estimations transform point data to a continuous surface:

$$\lambda(z)= \sum_{i=1}^{n} \frac{1}{\pi \tau^2} k(\frac{d_{iz}}{\tau})y_i$$
Where, 

$\lambda$(z) = density at location z; 

$\tau$ is bandwidth distance;

$k$ is the kernel function, typically a function of the ratio of $d_{iz}$ to $\tau$;

$d_{iz}$ is the distance from event $i$ to location $z$.

Using the Quartic function: 

$$\lambda(z)= \sum_{i=1}^{n} \frac{1}{\pi \tau^2}(\frac{3}{\pi}(1-\frac{d_{iz}^2}{\tau^2}))y_i$$

- We use the [density.ppp](https://rdrr.io/github/spatstat/spatstat/man/density.ppp.html) function from the [spatstat](https://cran.r-project.org/web/packages/spatstat/index.html ) package
  - We test 4 bandwidth ($\tau$) values: 50m,100m,250m and 500m
  - We compare hotspots derived by taking the top 10%, 5%, 1% and 0.1% of raster cells
  - We keep the cell size constant at 10m x 10m 


### Hotspots Interactive Map

- Hotspots displayed as a contour map
- Each bandwidth layer can be toggled on and off by clicking the layers icon, below the zoom buttons, on the left hand side

```{r,message=FALSE,warning=FALSE}

# study_area_sp <- as(st_as_sfc(st_bbox(study_area)),"Spatial")
study_area_sp <- as(study_area,"Spatial")
inc_sp <- as(incidents,"Spatial")
inc_ppp <- as(inc_sp,"ppp") #create ppp object from bikemaps points
Window(inc_ppp) <- as(study_area_sp,"owin") #define study extent in ppp object

bandwidth <- c(50,100,150,250)

inc_planar <- lapply(1:length(bandwidth),function(x){
  density(inc_ppp, 
          kernel = "quartic",
          sigma = bandwidth[x],
          eps = 10,
          positive=TRUE)
  }) #kernel density images

inc_planar_r <- lapply(inc_planar,function(x) raster(x,crs = CRS('+init=EPSG:26910'))) #convert to raster file
names(inc_planar_r) <-  bandwidth#convert to raster file
incident_hotspots <- lapply(inc_planar_r, function(x) detect_planar_hotspot(x,percentiles = c(0.90,0.95,0.99,0.999)))

#interactive plot
mapview(incident_hotspots,alpha.regions=0.25)
```

### Hotspots Map by Threshold and Bandwidth

```{r,fig.align = "center",fig.cap="Hotspots for incidents reported to BikeMaps.org by bandwidth size, and percentile threshold"}
#Plot in ggplot2
incident_hotspots_cmbnd <- do.call(what = sf:::rbind.sf, args = incident_hotspots) %>% 
  mutate(bw = row.names(.)) %>% 
  separate(col = bw,into = c("bw",NA)) %>% 
  mutate(bw=factor(bw,levels = as.character(bandwidth)))

pal <- RColorBrewer::brewer.pal(6,name = "Reds")[-(1:2)]

ggplot()+
  geom_sf(data=st_union(study_area),fill="snow1") +
  geom_sf(data=incident_hotspots_cmbnd,aes(fill=layer,colour=layer)) +
  scale_colour_manual(values = pal) + 
  scale_fill_manual(values = pal) + 
  facet_grid(bw~layer) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.minor = element_blank() ,
        panel.grid.major = element_blank() ,
        axis.title = element_blank(),legend.position = "none") +
  coord_sf() + 
  ggtitle("Incident Kernel Density: By Threshold (Columns) & Bandwidth (rows)")
```

### Quantifying differences

```{r}
hs <- lapply(1:nrow(incident_hotspots_cmbnd),
             function(x) st_sf(st_cast(st_union(incident_hotspots_cmbnd[x,]), "POLYGON")) %>% 
               mutate(area= st_area(.))) 

qe <- tibble(threshold = incident_hotspots_cmbnd$layer,
       bandwidth = incident_hotspots_cmbnd$bw,
       m2 = sapply(hs,function(x){sum(x$area)}),
       avg_m2 = sapply(hs,function(x){mean(x$area)}),
       n = sapply(hs,FUN = nrow))

qe %>% 
  select(threshold,bandwidth,n) %>% 
  pivot_wider(names_from = bandwidth,values_from = n) %>% 
  knitr::kable()
```

- Number of hotspots by threshold and bandwidth
  - Larger bandwidth results in less distinct hotspots (e.g instead of many distinct hotspots, few larger hotspots)
  - Higher threshold results in fewer hotspots (generally) but this difference is much more pronounced for smaller bandwidths than for larger ones
    - e.g. 6 distinct hotspots for a 500m bandwidth with 90% threshold, compared to 1 with 99.9% threshold. Compare this with 2302 distinct hotspots for a 50m bandwidth with a 90% threshold and 13 for a 99.9% threshold 

```{r}

pal <- RColorBrewer::brewer.pal(6,name = "Purples")[-(1:2)]

qe %>% 
ggplot(aes(x=bandwidth,y=avg_m2,colour=threshold,group=threshold)) + 
  geom_point() + 
  scale_colour_manual(values = pal) + 
  geom_line() + theme_minimal()

``` 

- Increasing the bandwidth results in larger and fewer hotspots
- Decreasing the bandwidth you an increased number of hotspots of a smaller size

# Network Kernel Density Surfaces

To produce density estimates on network:

$$\lambda(z)= \sum_{i=1}^{n} \frac{1}{\tau} k(\frac{d_{iz}}{\tau})y_i$$

Using the Quartic function: 

$$\lambda(z)= \sum_{i=1}^{n} \frac{1}{\tau}(\frac{3}{\pi}(1-\frac{d_{iz}^2}{\tau^2}))y_i$$

- I created function in R to implement network based KDE from algorithm outlined in [Xie & Yan (2008)](https://www.sciencedirect.com/science/article/pii/S0198971508000318)
  - The network KDE is a a 1-D version of the planar kernel density estimator, with $\tau$ (bandwidth) based on network distances, rather than Euclidean distances
  - Network split into "lixels" (1-D version of pixels)
  - We test 4 bandwidth ($\tau$) values: 50m,100m,250m and 500m
  - We compare hotspots derived by taking the top 10%, 5%, 1% and 0.1% of lixel density values
  - We keep the lixel size constant at 10m.


### Hotspots Interactive Map

- Hotspots displayed as a contour map
- Each bandwidth layer can be toggled on and off by clicking the layers icon, below the zoom buttons, on the left hand side

```{r}

inc_network_kde <- read_sf(paste0(getwd(),"/output/bikemaps_kde_10m_lixel_50m_250m_bw.gpkg"))

cols_vec <- c("kde_bw_50","kde_bw_100","kde_bw_150","kde_bw_250")

inc_network_hotspots <- lapply(cols_vec, function(x) detect_network_hotspot(inc_network_kde,x,percentiles = c(0.90,0.95,0.99,0.999)))
names(inc_network_hotspots) <- bandwidth

#interactive plot
mapview(inc_network_hotspots,zcol=list("thresholds","thresholds","thresholds","thresholds"))
```

### Hotspots Map by Threshold and Bandwidth

```{r}
#plot in ggplot
pal <- RColorBrewer::brewer.pal(6,name = "Reds")[-(1:2)]

inc_network_hotspots_cmbnd <- do.call(what = sf:::rbind.sf, args = inc_network_hotspots) %>% 
  mutate(bw = row.names(.)) %>% 
  separate(col = bw,into = c("bw",NA)) %>% 
  mutate(bw = factor(bw, levels = as.character(bandwidth)))

ggplot()+
  geom_sf(data=st_union(study_area),fill="snow1") +
  geom_sf(data=inc_network_hotspots_cmbnd,aes(fill=thresholds,colour=thresholds)) +
  scale_colour_manual(values = pal) + 
  scale_fill_manual(values = pal) + 
  facet_grid(bw~thresholds) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.minor = element_blank() ,
        panel.grid.major = element_blank() ,
        axis.title = element_blank(),legend.position = "none") +
  coord_sf() + 
  ggtitle("Network Kernel Density: Percentile Threshold (Columns), Bandwidth (rows)")

```


### Quantifying Differences

```{r}

hs <- lapply(1:nrow(inc_network_hotspots_cmbnd),
             function(x) st_sf(
               st_cast(
                 st_union(
                   st_buffer(
                     inc_network_hotspots_cmbnd[x,],dist=0.5
                   )
                 ), "POLYGON"
               )
             ) %>% mutate(area= st_area(.))) 

qe <- tibble(threshold = inc_network_hotspots_cmbnd$thresholds,
       bandwidth = inc_network_hotspots_cmbnd$bw,
       m = sapply(hs,function(x){sum(x$area)}),
       avg_m = sapply(hs,function(x){mean(x$area)}),
       n = sapply(hs,FUN = nrow))

qe %>% 
  select(threshold,bandwidth,n) %>% 
  pivot_wider(names_from = bandwidth,values_from = n) %>% 
  knitr::kable()

```

- Number of hotspots by threshold and bandwidth
  - Larger bandwidth results in less distinct hotspots (e.g instead of many distinct hotspots, few larger hotspots)
  - Higher threshold results in fewer hotspots (generally) but this difference is much more pronounced for smaller bandwidths than for larger ones
    - e.g. 6 distinct hotspots for a 500m bandwidth with 90% threshold, compared to 1 with 99.9% threshold. Compare this with 2302 distinct hotspots for a 50m bandwidth with a 90% threshold and 13 for a 99.9% threshold 

```{r}

pal <- RColorBrewer::brewer.pal(6,name = "Purples")[-(1:2)]

qe %>% 
ggplot(aes(x=bandwidth,y=avg_m,colour=threshold,group=threshold)) + 
  geom_point() + 
  scale_colour_manual(values = pal) + 
  geom_line() + theme_minimal()

``` 

- Increasing the bandwidth results in larger and fewer hotspots
- Decreasing the bandwidth you an increased number of hotspots of a smaller size

#Integrating Exposure

- Previous hotspots represent areas of high crash burden
- To estimate risk we need to integrate information on exposure 
- We take a simple approach to estimating crash risk density by dividing a kernel density estimate of crashes over a kernel density estimate of exposure: 
$$\lambda(z_{r}) = \frac{\lambda(z_{events})}{\lambda(z_{exposure})}$$

- This can be applied to either planar or network based density estimates 

### Planar Kernel Density

- To estimate a planar exposure surface we use [density.psp](https://rdrr.io/cran/spatstat/man/density.psp.html) from spatstat package
  - Weighted by strava counts
  
- We compare hotspots derived by taking the top 10%, 5%, 1% and 0.1% of pixel density values


```{r,warning=FALSE}

# Create planar density surface for exposure
network_sp <- as(network,"Spatial")
exp_psp <- as(network_sp[,"CAADB"],"psp")
Window(exp_psp) <- as(study_area_sp,"owin") #define study extent in ppp object

exp_planar <- lapply(1:length(bandwidth),function(x){
  density.psp(exp_psp,
          kernel = "quartic",
          sigma = bandwidth[x],
          eps = 10,
          weights = exp_psp$marks)
  }) #kernel density images


#recale exposure density to be positive values
exp_planar_r <- lapply(exp_planar,function(x) raster(x,crs = CRS('+init=EPSG:26910'))) #convert to raster file
for(i in 1:length(exp_planar_r)) {
  if(any(exp_planar_r[[i]]@data@values<0,na.rm=TRUE)){
    exp_planar_r[[i]]@data@values <- scales::rescale(exp_planar_r[[i]]@data@values,
                                                     to=c(0.000001,max(exp_planar_r[[i]]@data@values,na.rm = TRUE)))
  }
  }

## Divide incident raster surface by exposure planar surface

risk_planar <- mapply(`/`,inc_planar_r,exp_planar_r)

#find risk hotspots
risk_hotspots <- lapply(risk_planar, function(x) detect_planar_hotspot(x,percentiles = c(0.90,0.95,0.99,0.999)))


#interactive plot
mapview(risk_hotspots,alpha.regions=0.25) 

#Plot in ggplot2
risk_hotspots_cmbnd <- do.call(what = sf:::rbind.sf, args = risk_hotspots) %>% 
  mutate(bw = row.names(.)) %>% 
  separate(col = bw,into = c("bw",NA)) %>% 
  mutate(bw=factor(bw,levels = as.character(bandwidth)))

pal <- RColorBrewer::brewer.pal(6,name = "Reds")[-(1:2)]

ggplot()+
  geom_sf(data=st_union(study_area),fill="snow1") +
  geom_sf(data=risk_hotspots_cmbnd,aes(fill=layer,colour=layer)) +
  scale_colour_manual(values = pal) + 
  scale_fill_manual(values = pal) + 
  facet_grid(bw~layer) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.minor = element_blank() ,
        panel.grid.major = element_blank() ,
        axis.title = element_blank(),legend.position = "none") +
  coord_sf() + 
  ggtitle("Risk Kernel Density: Percentile Threshold (Columns), Bandwidth (rows)")


```

### Network Kernel Density

- We compare hotspots derived by taking the top 10%, 5%, 1% and 0.1% of lixel density values

```{r}
exposure_network_kde <- read_sf(paste0(getwd(),"/output/strava_kde_10m_lixel_50m_500m_bw.gpkg"))

risk_network_kde <- left_join(inc_network_kde,
                              exposure_network_kde %>% st_drop_geometry(),
                              by = c("ID","LIXID","length"))
risk_network_kde <- risk_network_kde %>% 
  mutate(risk_kde_bw_50 = kde_bw_50.x/kde_bw_50.y,
         risk_kde_bw_100 = kde_bw_100.x/kde_bw_100.y,
         risk_kde_bw_250 = kde_bw_250.x/kde_bw_250.y,
         risk_kde_bw_500 = kde_bw_500.x/kde_bw_500.y
  )


cols_vec <- c("risk_kde_bw_50","risk_kde_bw_100","risk_kde_bw_250","risk_kde_bw_500")

risk_network_hotspots <- lapply(cols_vec, function(x) detect_network_hotspot(risk_network_kde,x,percentiles = c(0.90,0.95,0.99,0.999)))
names(risk_network_hotspots) <- bandwidth

#interactive plot
mapview(risk_network_hotspots,zcol=list("thresholds","thresholds","thresholds","thresholds"))

```


# Main Takeaways

- Bandwidth very important (obvious a-priori) and will reflect different spatial processes for a given scale
  - Larger bandwidth results in larger hotspots that likely reflect neighbourhood level processes such as land use, connectivity that lead to greater density
  - Smaller bandwidth results in smaller hotspots, more reflective of site specific conditions such as geometric design etc. 

- Threshold more of a measure of resources. How many sites would a planner wish to investigate further, how much resources are needed? Choose to balance the number of hotspots versus available resources, set higher threshold to isolate fewer hotspots. 

- Planar KDE less precise, especially at larger bandwidths. E.g. the hotspot for 500m planar KDE bandwidth much further east in downtown for 99.9 percentile compared to the equivalent network KDE which has the hotspot much closer to Johnson street bridge. Planar KDE assumes risk is constant over space, doesn't take into account differences in network density downtown.  

- Integrating exposure requires spatially comprehensive exposure data - bias corrected fitness app counts are one of the Only means of obtaining reasonable estimates of bicycling exposure data for an entire study area.
  - as an aside this would be a great selling point for any kind of future consulting the bikemaps team could do in the future: 
    - 1) We can generate great spatially representative exposure data
    - 2) This will enable us to identify spatial hotspots of _risk_ in addition to burden, in a simple straight forward manner that would not be possible without the bias-corrected exposure data


# Moving forward...

Options for further comparisons:

1. Use LISA (Moran's I or Getis-Ord) to identify density hotspots amongst lixels with monte carlo simulation for significance (see [Xie and Yan](https://www.sciencedirect.com/science/article/pii/S0966692313000860)
  - This would be a way of identifying high-high clusters for kernel density based on psuedo significance
  - I have gotten this to work but need time to improve computation times to make it feasible. Confident I can do it but a matter of balancing the time and value added? 






